---
title: "LOCUS & SATS Project"
author: "Matt Dunham"
date: "7/9/2020"
output: html_document
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Packages}

library(GPArotation)
library(nFactors)
library(psych)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lavaan)
library(semPlot)
library(kableExtra)
library(magrittr)
library(RColorBrewer)
library(readxl)
library(moments)
library(reshape)
library(resemble)
library(likert)
library(broom)
library(grid)
library(patchwork)
library(kableExtra)
library(matrixStats)
library(networkD3)
library(fs)
library(plyr)
```

# Reading in the Data

Throughout the code, two data sets will be used - one for the SATS data and one for the LOCUS data. Procedures, such as reading in and cleaning the data, are different for the two data sets. 

## SATS Data

The SATS data was initially downloaded from Qualtrics as a numeric excel file. Excel files are generally easier to work with and require little changes in code compared to the traditional cvs file. After being downloaded, the second row of each data set was deleted as it was a duplicate column header and caused some issues when reading in the data. No other changes to the data were made prior to reading it in.

```{r Reading in SATS Data}

semesters <- c("FA17", "FA18", "FA19", "FA20", "SP17", "SP18", "SP19", "SP20") ### Creating a list of each potential semester the data could be from

status <- c("PRE", "POST") ### List of either pre or post status

class <- c(100, 250) ### list of potential classes (can only be STAT)

file_names_sats <- dir(path="C:/Users/Matth/OneDrive/Desktop/LOCUS_SOMAS/SATS Data") ### Reading in the names of all the files from a seperate folder

data_sats <- data.frame() ### Creating an empty data frame to fill with files

for (i in 1:length(file_names_sats)) { ### Loop the amount of times there are files in the folder
    
    name <- file_names_sats[i] ### Storing the name of the data file 

    temp_data_sats <- read_excel(file_names_sats[i]) ### Reading in a data file into a temporary data set
    
    for (j in 1:length(semesters)) { ### Checking the title of the file to see which semester it belongs to
      
        if (grepl(semesters[j], name) == 1) { ### If the semester matches
            temp_data_sats$semester = semesters[j] ### Create a column and fill it in with that semester
        }
    }
    
    for (j in 1:length(status)) { ### Checking the title of the file to see if it is pre or post
      
        if (grepl(status[j], name) == 1) { ### If the title matches
            temp_data_sats$pre_post = status[j] ### Add "PRE" or "POST" to a new column
        }
    }
    
    for (j in 1:length(class)) { ### Checking the totle of the file to see which class the data is rom
      
        if (grepl(class[j], name) == 1) { ### Once the class is identified
            temp_data_sats$class = class[j] ### Add a column with the number (classes are always STAT)
        }
    }
    

    data_sats <- bind_rows(data_sats, temp_data_sats) ### bind the first data fild with this new one, combining the two
    
    ### Each loop will add an aditional data file to the data frame, eventually adding all the files within the folder.
} %>%
  mutate(A4 = 8 - A4) %>%
  mutate(C5 = 8 - C5) %>%
  mutate(V7 = 8 - V7) %>%
  mutate(D8 = 8 - D8) %>%
  mutate(C11 = 8 - C11) %>%
  mutate(V13 = 8 - V13) %>%
  mutate(A15 = 8 - A15) %>%
  mutate(V16 = 8 - V16) %>%
  mutate(A18 = 8 - A18) %>%
  mutate(V21 = 8 - V21) %>%
  mutate(D24 = 8 - D24) %>%
  mutate(V25 = 8 - V25) %>%
  mutate(C26 = 8 - C26) %>%
  mutate(A28 = 8 - A28) %>%
  mutate(D30 = 8 - D30) %>%
  mutate(V33 = 8 - V33) %>%
  mutate(D34 = 8 - D34) %>%
  mutate(C35 = 8 - C35) %>%
  mutate(D36 = 8 - D36) 

sats_k <- length(file_names_sats) ### Calculating number of files in the folder

sats_n <- length(data_sats$StartDate) ### Calculating sample size 


```

We have a total of **`r sats_k`** files and **`r sats_n`** observations.

## LOCUS Data

LOCUS data was downloaded from the LOCUS website. The data downloaded was "Test Taker Raw" which contains students responses (we can calculate scores from this). Every file was converted to an excel file as I was having trouble reading it as a csv (this may be due to the comma seperated values). I did this manually, but I will attempt to find a way to have this done more efficiently.

```{r Reading in LOCUS Data, message=FALSE, warning=FALSE}

file_names_locus <- dir(path="C:/Users/Matth/OneDrive/Desktop/LOCUS_SOMAS/LOCUS Data") ### Reading in the names of all the files from a seperate folder

data_locus <- data.frame() ### Creating an empty data frame to fill with files

for (i in 1:length(file_names_locus)) {  ### Loop the amount of times there are files in the folder
    
    name <- file_names_locus[i] ### Storing the name of the data file 

    temp_data_locus <- read_excel(file_names_locus[i]) ### Reading in a data file into a temporary data set
    
    for (j in 1:length(semesters)) { ### Checking the title of the file to see which semester it belongs to
      
        if (grepl(semesters[j], name) == 1) { ### If the semester matches
            temp_data_locus$semester = semesters[j] ### Create a column and fill it in with that semester
        }
    }
    
    for (j in 1:length(status)) { ### Checking the title of the file to see if it is pre or post
      
        if (grepl(status[j], name) == 1) { ### If the title matches
            temp_data_locus$pre_post = status[j] ### Add "PRE" or "POST" to a new column
        }
    }
    
    for (j in 1:length(class)) { ### Checking the totle of the file to see which class the data is rom
      
        if (grepl(class[j], name) == 1) { ### Once the class is identified
            temp_data_locus$class = class[j] ### Add a column with the number (classes are always STAT)
        }
    }
    
    data_locus <- bind_rows(data_locus, temp_data_locus) ### bind the first data fild with this new one, combining the two
    
    ### Each loop will add an aditional data file to the data frame, eventually adding all the files within the folder.
}

locus_k <- length(file_names_locus) ### Calculating number of files in the folder

locus_n <- length(data_locus$Score) ### Calculating sample size 

```

We have a total of **`r locus_k`** files and **`r locus_n`** observations.

# Cleaning Data

There are various things we need to do to our data before it is ready for analysis. Each file is seperate, but the general idea is to move from our two seperated filed to a single merged file with Student name - Sats pre - Sats post - Locus pre - Locus post. The process of getting to that will be seperate for each data file, but we will eventually merge the files together.

## SATS Data

First off, we need to merge our name columns together in order to be able to match our students together. To do this, we will simply merge our columns together that contain the students names, as well as make the names all lower case to create a consistent format.

```{r Cleaning SATS Data}

sats <- data_sats %>%
  mutate(fullname = paste(FirstName, LastName, seq=' ')) %>% ### Combining first and last name column
  select(E1:D36, semester, pre_post, class, fullname) %>% ### Selecting item responses, semester, pre_post status, class and student name
  mutate(na = rowSums(is.na(.[1:36]))) %>% ### Determining how many NA values in each row and creating a new column with that value for each row
  filter(na == 0) %>% ### Removing any row with any NA values (potentially change this?)
  select(-na) %>% ### Remove the column with na value
  mutate(SC0 = rowSums(.[1:36])) %>% ### Calculating final scores
  select(SC0, fullname, pre_post, semester, class) %>% ### selecting desired columns
  mutate(fullname = str_to_lower(fullname)) %>% ### Removing spacing and weird characters to keep a consistent formatt
  mutate(fullname = str_replace_all(fullname, fixed(" "), "")) %>%
  mutate(fullname = str_replace_all(fullname, fixed("-"), "")) %>%
  mutate(fullname = str_replace_all(fullname, fixed("'"), "")) 

sats_scores <- sats %>%
  mutate(sats_pre = ifelse(pre_post %in% "PRE", SC0, '')) %>% ### Creating a new column or pre scores and filling it with the SATS score if the columns contains PRE
  mutate(sats_post = ifelse(pre_post %in% "POST", SC0, '')) %>% ### Doing the same for post
  select(fullname, semester, sats_pre, sats_post, class) %>% ### selecting names, senester, and pre/post score
  group_by(fullname, semester, class) %>% ### grouping by names and semester
  summarise_all(funs(trimws(paste(., collapse = '')))) ### If the names and semesters and the same, the row is collapsed and since the scores columns should contain nothing depending on the pre/post status, the rows should collapse cleanly and line up.

```

## LOCUS Data

```{r Cleaning LOCUS Data}

### Same process for LOCUS
locus <- data_locus %>%
  mutate(firstname = sapply(strsplit(data_locus$`Test Taker Identifier`, " "), head, 1)) %>% ### Taking first name column
  mutate(lastname = sapply(strsplit(data_locus$`Test Taker Identifier`, " "), tail, 1)) %>% ### Taking last name column
  mutate(fullname = paste(firstname, lastname, seq=' ')) %>% ## Putting them together with a single space inbetween
  select(Score:'Question 23...27', semester, pre_post, class, fullname) %>% ### Selecting columns of interest
  mutate(na = rowSums(is.na(.[2:24]))) %>% ### Determining how many NA values in each row and creating a new column with that value for each row
  filter(na == 0) %>% ### Removing any row with any NA values (potentially change this?)
  select(-na) %>% ### Remove the column with na value 
  select(Score, pre_post, fullname, semester, class) %>% ### Select desired columns
  ### Removing spacing and weird symbols from names to keep format consisent.
  mutate(fullname = str_to_lower(fullname)) %>%
  mutate(fullname = str_replace_all(fullname, fixed(" "), "")) %>%
  mutate(fullname = str_replace_all(fullname, fixed("-"), "")) %>%
  mutate(fullname = str_replace_all(fullname, fixed("'"), "")) 

locus_scores <- locus %>%
  mutate(locus_pre = ifelse(pre_post %in% "PRE", Score, '')) %>% ### Creating a column that fills in with the students score if the student took the PRE test and with nothing if they took the POST test
  mutate(locus_post = ifelse(pre_post %in% "POST", Score, '')) %>% ### Doing the same but for POST
  select(fullname, semester, locus_pre, locus_post, class) %>% ### Selecting names, class, semester and pre_post status
  group_by(fullname, semester, class) %>% ### Grouping by name, semester and class
  summarise_all(funs(trimws(paste(., collapse = '')))) ### Collapsing rows together together if they have the same name, semester and class, which will line up their pre_post scores.

```

To do tomorrow: 

* Current issue is figuring out whats up with when students take the assessments multiple times.

* Getting the final columns under way

* Combining the names together

```{r Merging Data Together}

### This merges our data files together

final_scores <- merge(sats_scores, locus_scores, on=c("fullname", "semester", "class")) ### Merging the sats and locus data together, combining rows if they have the name full name, semester and class

final_scores_cleaned <- final_scores %>%
  mutate_all(~ifelse(. %in% c("N/A", "null", ""), NA, .)) %>% ### Removing NA values
  na.omit()

```

# Summary Statistics

We can look a specific item responses and group them into constructs previously identified in SATS and represent them various ways. We are really interested in the pre and post differences between scores.

## SATS Data

We can produce overlapping histograms with a mean line to see differences across a semester via pre/post averages. This will be done for all the SATS constructs, as well as for the LOCUS scores.

### Effort

```{r SATS items}

sats_effort <- data_sats %>%
  select(E1, E2, E14, E27, pre_post) %>% 
  mutate_all(~ifelse(. %in% c("N/A", "null", ""), NA, .)) %>%
  na.omit()

sats_effort_mean <- sats_effort %>%
  select(E1:E27) %>%
  mutate(means = rowMeans(.)) %>% ### Calculating means for each row (or student in this case) so that we can plot a histogram of the averages in the constructs.
  mutate(pre_post = sats_effort$pre_post) 

sats_effort_post <- subset(sats_effort_mean, pre_post == "POST")
sats_effort_pre <- subset(sats_effort_mean, pre_post == "PRE")

ggplot(sats_effort_mean, aes(x=means)) +
  geom_histogram(data=subset(sats_effort_mean, pre_post == "POST"), fill = "red", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_effort_post$means)) +
  geom_histogram(data=subset(sats_effort_mean, pre_post == "PRE"), fill = "green", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_effort_pre$means), linetype = "dashed") +
  scale_x_continuous(name="effort", limits = c(0, 8))
  
```

### Interest 

```{r}
sats_interest <- data_sats %>%
  select(I12, I20, I23, I29, pre_post) %>%
  mutate_all(~ifelse(. %in% c("N/A", "null", ""), NA, .)) %>%
  na.omit()

sats_interest_mean <- sats_interest %>%
  select(I12:I29) %>%
  mutate(means = rowMeans(.)) %>% ### Calculating means for each row (or student in this case) so that we can plot a histogram of the averages in the constructs.
  mutate(pre_post = sats_interest$pre_post) 

sats_interest_post <- subset(sats_interest_mean, pre_post == "POST")
sats_interest_pre <- subset(sats_interest_mean, pre_post == "PRE")

ggplot(sats_interest_mean, aes(x=means)) +
  geom_histogram(data=subset(sats_interest_mean, pre_post == "POST"), fill = "red", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_interest_post$means)) +
  geom_histogram(data=subset(sats_interest_mean, pre_post == "PRE"), fill = "green", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_interest_pre$means), linetype = "dashed") +
  scale_x_continuous(name="interest", limits = c(0, 8))

```

### Difficulty

```{r}
sats_difficulty <- data_sats %>%
  select(D6, D8, D22, D24, D30, D34, D36, pre_post) %>%
  mutate_all(~ifelse(. %in% c("N/A", "null", ""), NA, .)) %>%
  na.omit()

sats_difficulty_mean <- sats_difficulty %>%
  select(D6:D36) %>%
  mutate(means = rowMeans(.)) %>% ### Calculating means for each row (or student in this case) so that we can plot a histogram of the averages in the constructs.
  mutate(pre_post = sats_difficulty$pre_post) 

sats_difficulty_post <- subset(sats_difficulty_mean, pre_post == "POST")
sats_difficulty_pre <- subset(sats_difficulty_mean, pre_post == "PRE")

ggplot(sats_difficulty, aes(x=means)) +
  geom_histogram(data=subset(sats_difficulty_mean, pre_post == "POST"), fill = "red", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_difficulty_post$means)) +
  geom_histogram(data=subset(sats_difficulty_mean, pre_post == "PRE"), fill = "green", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_difficulty_pre$means), linetype = "dashed") +
  scale_x_continuous(name="difficulty", limits = c(0, 8))
```

### Value

```{r}
sats_value <- data_sats %>%
  select(V7, V9, V10, V13, V16, V17, V21, V25, V33, pre_post) %>%
  mutate_all(~ifelse(. %in% c("N/A", "null", ""), NA, .)) %>%
  na.omit()

sats_value_mean <- sats_value %>%
  select(V7:V33) %>%
  mutate(means = rowMeans(.)) %>% ### Calculating means for each row (or student in this case) so that we can plot a histogram of the averages in the constructs.
  mutate(pre_post = sats_value$pre_post) 

sats_value_post <- subset(sats_value_mean, pre_post == "POST")
sats_value_pre <- subset(sats_value_mean, pre_post == "PRE")

ggplot(sats_value, aes(x=means)) +
  geom_histogram(data=subset(sats_value_mean, pre_post == "POST"), fill = "red", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_value_post$means)) +
  geom_histogram(data=subset(sats_value_mean, pre_post == "PRE"), fill = "green", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_value_pre$means), linetype = "dashed") +
  scale_x_continuous(name="value", limits = c(0, 8))
```

### Cognitive Competence

```{r}
sats_cogcomp <- data_sats %>%
  select(C5, C11, C26, C31, C32, C35, pre_post) %>%
  mutate_all(~ifelse(. %in% c("N/A", "null", ""), NA, .)) %>%
  na.omit()

sats_cogcomp_mean <- sats_cogcomp %>%
  select(C5:C35) %>%
  mutate(means = rowMeans(.)) %>% ### Calculating means for each row (or student in this case) so that we can plot a histogram of the averages in the constructs.
  mutate(pre_post = sats_cogcomp$pre_post) 

sats_cogcomp_post <- subset(sats_cogcomp_mean, pre_post == "POST")
sats_cogcomp_pre <- subset(sats_cogcomp_mean, pre_post == "PRE")

ggplot(sats_cogcomp, aes(x=means)) +
  geom_histogram(data=subset(sats_cogcomp_mean, pre_post == "POST"), fill = "red", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_cogcomp_post$means)) +
  geom_histogram(data=subset(sats_cogcomp_mean, pre_post == "PRE"), fill = "green", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_cogcomp_pre$means), linetype = "dashed") +
  scale_x_continuous(name="cogcomp", limits = c(0, 8))
```

### Affect

```{r}
sats_affect <- data_sats %>%
  select(A3, A4, A15, A18, A19, A28, pre_post) %>%
  mutate_all(~ifelse(. %in% c("N/A", "null", ""), NA, .)) %>%
  na.omit()

sats_affect_mean <- sats_affect %>%
  select(A3:A28) %>%
  mutate(means = rowMeans(.)) %>% ### Calculating means for each row (or student in this case) so that we can plot a histogram of the averages in the constructs.
  mutate(pre_post = sats_affect$pre_post) 

sats_affect_post <- subset(sats_affect_mean, pre_post == "POST")
sats_affect_pre <- subset(sats_affect_mean, pre_post == "PRE")

ggplot(sats_affect_mean, aes(x=means)) +
  geom_histogram(data=subset(sats_affect_mean, pre_post == "POST"), fill = "red", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_affect_post$means)) +
  geom_histogram(data=subset(sats_affect_mean, pre_post == "PRE"), fill = "green", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_affect_pre$means), linetype = "dashed") +
  scale_x_continuous(name="affect", limits = c(0, 8))
```
### Scores

```{r}
sats_responses <- sats %>%
  select(SC0, pre_post)

sats_responses_post <- subset(sats_responses, pre_post == "POST")
sats_responses_pre <- subset(sats_responses, pre_post == "PRE")

ggplot(sats_responses, aes(x=SC0)) +
  geom_histogram(data=subset(sats_responses, pre_post == "POST"), fill = "red", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_responses_post$SC0)) +
  geom_histogram(data=subset(sats_responses, pre_post == "PRE"), fill = "green", alpha = 0.2) +
  geom_vline(xintercept = mean(sats_responses_pre$SC0), linetype = "dashed") +
  scale_x_continuous(name="scores")
```

## LOCUS Data

### Scores

```{r}
locus_responses <- data_locus %>%
  select(Score, pre_post)

locus_responses_post <- subset(locus_responses, pre_post == "POST")
locus_responses_pre <- subset(locus_responses, pre_post == "PRE")

ggplot(locus_responses, aes(x=Score)) +
  geom_histogram(data=subset(locus_responses, pre_post == "POST"), fill = "red", alpha = 0.2) +
  geom_vline(xintercept = mean(locus_responses_post$Score)) +
  geom_histogram(data=subset(locus_responses, pre_post == "PRE"), fill = "green", alpha = 0.2) +
  geom_vline(xintercept = mean(locus_responses_pre$Score), linetype = "dashed") +
  scale_x_continuous(name="scores")
```